---
title: "Pennsylvania SDWIS Violations"
author: "Walker Grimshaw"
date: "2/4/2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

library(tidyverse)
library(readxl)
library(lubridate)
library(plotly)
```

## Read in SDWIS Violation Data

```{r}
## Raw violations data
SDWIS_viol_raw <- read.csv("./DATA/RAW/PAPWSID_SDWIS_Violation.csv")

## Raw Enforcement Action data
SDWIS_Enforcement_Raw <- read.csv("./DATA/RAW/PAPWSID_SDWIS_Enforcement_Action.csv")

## Raw violation enforcement assoc
SDWIS_Assoc <- read.csv("./DATA/RAW/PAPWSID_SDWIS_Violation_Enf_Assoc.csv")

## Pennvest data
Pennvest <- read_excel("./DATA/RAW/PA_Pennvest_Loans_and_Grants.xlsx")
```

## Clean Violations Data

```{r}
## clean the whole violations database
SDWIS_viol_clean <- SDWIS_viol_raw %>%
  ## remove "VIOLATION." from the name of each column
  rename_all(funs(gsub("VIOLATION.", "", names(SDWIS_viol_raw)))) %>%
  ## remove indexing column and last columns that is just NAs
  select(-X.1,-X) %>%
  ## filter for only PA systems
  filter(PRIMACY_AGENCY_CODE == "PA") %>%
  ## change start date and end date to dates
  mutate(COMPL_PER_BEGIN_DATE = as.Date(COMPL_PER_BEGIN_DATE, format = "%d-%b-%y"),
         COMPL_PER_END_DATE = as.Date(COMPL_PER_END_DATE, format = "%d-%b-%y"),
         RTC_DATE = as.Date(RTC_DATE, format = "%d-%b-%y")) %>%
  ## add year column
  mutate(Year = year(COMPL_PER_BEGIN_DATE),
         ## Add Severity Index Column
         Severity_Index = case_when(IS_HEALTH_BASED_IND == "Y" ~ 10,
                                    IS_MAJOR_VIOL_IND == "Y" ~ 1,
                                    TRUE ~ 0))

```

## Clean Enforcement Action Data

```{r}
SDWIS_Enforcement_Clean <- SDWIS_Enforcement_Raw %>%
  ## remove "Enforcement action." from the beginning of each column name
  rename_all(funs(gsub("ENFORCEMENT_ACTION.", "", names(SDWIS_Enforcement_Raw)))) %>%
  ## remove first and last columns with no information
  select(-X.1, -X) %>%
  ## filter for only PA systems
  filter(grepl("PA", PWSID)) %>%
  ## change Enforcement Date to be a date
  mutate(ENFORCEMENT_DATE = as.Date(ENFORCEMENT_DATE, format = "%d-%b-%y"))
  
```

## Cities of Interest

```{r}
## create PWSID and PWS_Name dataframe from Pennvest data
Struggling_Names <- Pennvest %>%
  select(PWSID, SystemName) %>%
  distinct()

## Shrinking Cities violations data
SDWIS_Struggling <- SDWIS_viol_clean %>%
  ## filter for only the systems we are looking at
  filter(PWSID %in% Struggling_Names$PWSID) %>%
  ## add system name column
  left_join(Struggling_Names, by = "PWSID") %>%
  left_join(SDWIS_Enforcement_Clean,
            by = c("PWSID" = "PWSID", "LATEST_ENFORCEMENT_ID" = "ENFORCEMENT_ID")) %>%
  ## Add the word enforcement to the type_code column
  mutate(ENFORCEMENT_TYPE_CODE = TYPE_CODE) %>%
  ## drop type_code
  select(-TYPE_CODE, -ORIGINATOR_CODE.y) %>%
  ## add column of time between violation and return to compliance
  mutate(Noncompliance_Time = difftime(COMPL_PER_BEGIN_DATE, RTC_DATE, units = "days"))

```

## Summary Tables

``` {r}
## Shrinking Cities Severity Index
Struggling_Severity <- SDWIS_Struggling %>%
  group_by(SystemName, IS_HEALTH_BASED_IND) %>%
  summarize(Severity_Index = sum(Severity_Index)) %>%
  ## add system_type column corresponsing to struggling
  mutate(System_Group = "Struggling")

## SDWIS Severity
SDWIS_Severity <- SDWIS_viol_clean %>%
  group_by(PWSID) %>%
  summarize(Severity_Index = sum(Severity_Index)) %>%
  ## add system_group corresponding to PA
  mutate(System_Group = "PA")

## Combined Severity with System_Group
Severity_Group <- bind_rows(Struggling_Severity, SDWIS_Severity)

## Total Violations by type
SDWIS_Type <- SDWIS_shrinking %>%
  group_by(PWSID, CATEGORY_CODE) %>%
  summarize(count = n(), SystemName = first(SystemName))

## Total Violations by Health
SDWIS_Health <- filter(SDWIS_viol_clean, IS_HEALTH_BASED_IND == "Y") %>%
  group_by(PWSID, Year) %>%
  summarize(count = n())

## Shrinking City Health Based Violations
SDWIS_Shrinking_Health <- filter(SDWIS_shrinking, IS_HEALTH_BASED_IND == "Y") %>%
  group_by(SystemName, Year) %>%
  summarize(count = n())

## Just Major Violations
SDWIS_Shrinking_Major <- SDWIS_shrinking %>%
  filter(IS_MAJOR_VIOL_IND == "Y") %>%
  mutate(RULE_CODE = as.factor(RULE_CODE))

## Annual Violations
SDWIS_Annual <- SDWIS_shrinking %>%
  group_by(PWSID, Year) %>%
  summarize(count = n(), SystemName = first(SystemName))

## Annual Health Based Violations
SDWIS_Annual_Health <- SDWIS_shrinking %>%
  filter(IS_HEALTH_BASED_IND == "Y" | IS_MAJOR_VIOL_IND == "Y")
```


## Total Violations

```{r}
## Violations by System and Enforcement Type
SDWIS_Enforcement_plot <- ggplot(SDWIS_shrinking,
                                 aes(x = SystemName, fill = ENFORCEMENT_TYPE_CODE)) +
  geom_bar()
ggplotly(SDWIS_Enforcement_plot)

## Category Code
SDWIS_Type_plot <- ggplot(SDWIS_Type, aes(x = SystemName, y = count, fill = CATEGORY_CODE)) +
  geom_col()
SDWIS_Type_plot
ggplotly(SDWIS_Type_plot)

## Over time
SDWIS_Annual_plot <- ggplot(SDWIS_Annual, aes(x = Year, y = count, color = SystemName)) +
  geom_point()
SDWIS_Annual_plot
ggplotly(SDWIS_Annual_plot)


```


## Health Based Violations

```{r}
## compare shrinking distribution to total distribution
Health_Comparison_Plot <- ggplot() +
  geom_boxplot(data = SDWIS_Health,aes(group = Year, y = count, x = Year)) +
  geom_point(data = SDWIS_Shrinking_Health,
               aes(color = SystemName, y = count, x = Year))
ggplotly(Health_Comparison_Plot)


## Is health based
SDWIS_Health_plot <- ggplot(SDWIS_Shrinking_Health,
                            aes(x = SystemName, fill = IS_HEALTH_BASED_IND)) +
  geom_bar()
SDWIS_Health_plot
ggplotly(SDWIS_Health_plot)


```

## Major Violations

```{r}
Major_Violations_plot <- ggplot(SDWIS_Shrinking_Major) +
  geom_bar(aes(x = SystemName)) +
  labs(title = "Major Violations")
ggplotly(Major_Violations_plot)
```

## Severity Index Plots

```{r}
Struggling_Severity_plot <- ggplot(Struggling_Severity,
                                  aes(x = SystemName, y = Severity_Index,
                                      fill = IS_HEALTH_BASED_IND)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank()) +
  labs(y = "Severity Index") +
  scale_fill_discrete(name = "Violation Type",
                      labels = c("Major Monitoring\nReporting", "Health Based")) +
  ## Horizontal line corresponding to the median severity index for PA
  geom_hline(yintercept = median(SDWIS_Severity$Severity_Index),
             linetype = "dashed", color = "red", size = 1) +
  ## Horizontal line corresponding to 75th percentile of PA Severity Index
  geom_hline(yintercept = quantile(SDWIS_Severity$Severity_Index, probs = 0.75),
             linetype = "dashed", color = "red", size = 1)
Struggling_Severity_plot

Severity_Distribution <- ggplot(Severity_Group, aes(x = System_Group, y = Severity_Index)) +
  geom_boxplot() +
  coord_cartesian(ylim = c(0, 200)) +
  labs(y = "Severity Index") +
  theme(axis.title.x = element_blank())
Severity_Distribution
```

There is no apparent significant difference in total severity index between the struggling cities we have identified and all systems in Pennsylvania. If anything, the many outliers in Pennsylvania would raise the mean to be higher than that for the struggling cities. The struggling cities with the greatest

## Write Water Quality Summary Table

```{r}
SDWIS_Violation_Summary <- SDWIS_Struggling %>%
  group_by(PWSID) %>%
  summarize(SystemName = first(SystemName), Violation_Severity_Index = sum(Severity_Index))

## write to file
write.csv(SDWIS_Violation_Summary, "./DATA/PROCESSED/SDWIS_Violation_Summary.csv",
          row.names = F)
```

